Amazonian Narratives in Presidential Discourse (1985 to 2020)

### 2021-11-09 11112021_Sposito_and_SilvaMuller_DRAFT.Rmd
#color-blind friendly palette

---
title: "Amazonian Narratives in Presidential Discourse (1985 to 2020)"
author:
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
date: "11th of November of 2021"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

@@

Preserving the Amazon has rarely been the top priority of Brazilian federal governments. Though deforestation rates began to steadily increase in 2012 (see Carvalho et al. 2019), after a decade of decline, they spiked in 2019 after Bolsonaro took power (INPE 2021). This is no surprise as the current federal government is weakening the foundations of the Brazilian environmental governance system through budgetary cuts, presidential appointments, and dissociating government from civil society (see Ferrante and Fearnside 2019). These factors contributed to record-high fires in the Amazon and unprecedented monthly rates of deforestation during the dry seasons of 2019, 2020, and 2021 (INPE 2021). Although policy discontinuities between Bolsonaro's Amazon agenda and previous governments have been documented by academic scholarship and specialized media (see Abessa, Famá, and Buruaem 2019; Spring 2021), we lack empirical accounts of how Amazonian discourses changed. Pokorny et. al (2021, p. 2208), for example, argue that Bolsonaro’s Amazonian discourses shifted drastically in relation to earlier governments, giving rise to a “populist neoliberal” phase of discourse and policy developments in the region. However, the authors rely on a few disconnect quotes from Bolsonaro about the Amazon to exemplify phases defined by policy discontinuities. How, then, have Amazonian presidential discourses changed in time?

@@

Amazonian narratives constitute discourses that connect the region to specific imaginaries of the future and past. We argue that they touch on shared meanings about the Amazon that are available to the speaker as part of the larger social history of the region and country. To identify a few possible narratives, we utilize environmental historiography about the Amazon. We build Amazonian narratives as ideal types, that is, unified analytical constructs portraying a ‘pure’ version of a phenomenon (Weber 1949). In that sense, they rarely exist in its purest form. It serves the purpose of allocating empirical observations within a range of possibilities. The ideal types also help us deal with statements that perform different narratives. We broadly understand political discourses as performances seeking to appeal to diverse audiences across multiple political stages to project a definition of the situation (Goffman 1956, p. 3). As political stages, audiences, and time change, so do narratives. As with broader political discourses, narratives can appear together, not necessarily in consistent ways in time or across stages. As we show in section 3, our methodological strategy utilizing allows us to allocate statements within one or more ideal types, providing a more sophisticated portrayal of political discourse.

@@

The developmentalist narrative advances the view that the Brazilian Amazon needs to be developed and modernized. This includes expanding the agricultural frontier through incentives, creating infrastructure as roads and dams, and fostering industries through tax-free zones. It has its origins in the policies of the 1930s that attempted to modernize the recently consolidated region (Hecht and Cockburn, 1990), but were extended throughout the 20th century. It also captures arguments of the 20th century that defended the integration of indigenous populations into the nation via assimilation and inclusion in economic life. This narrative employs imaginaries of underdevelopment and economic needs in the Amazon while, at the same time, focus on the unexplored economic potential of the region and forest.

*Conservationism*

@@

# 4 Exploratory analysis

@@

# Remove stop words and punctuation
Amazon_speeches$text <- tm::removePunctuation(Amazon_speeches$text)
Amazon_speeches$text <- tm::removeWords(Amazon_speeches$text,
                                        c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre", "roraima",
                                          "manaus", "rondônia", "ministros", "luiz", "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "rio grande do sul", "rio grande do norte", "sao paulo", "então", "pouco", "pode", "sul",
                                          "norte", "mato grosso","outro", "onde", "discurso", "parte", "anos", "cada", "todo", 
                                          "porto velho", "visite site", "site secretaria", "minas gerais", "lula silva", 
                                          "inácio lula", "amapa", "bahia", "é", "lá", "et", "in", "httpwww", "vez", "vezes", "entao", "coisa", "ainda", "sobre", "apenas", "coisa",
                                          "jose", "tudo", "toda"))

@@

```{r location, echo=FALSE, warning=FALSE, message=FALSE}
# Location was added with poldis::extraxt_location(), still needs to be improved...
# To see the variable summary, please run: `summary(as.factor(Amazon_speeches$location))`
# Let's code the Amazonian states, the other Brazilian states, and speeches delivered outside Brazil.
Amazon_lc <- mutate(Amazon_speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian_States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non_Amazonian_States", grepl("NA", location) ~ "Non_Identified", !grepl("Amazonian_States|Non_Amazonian_States|Non_Identified", location) ~ "International"))
# Just a comparison with the proportion of speeches per stage on the broader dataset
stages <- mutate(BR_Presidential_Speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non Amazonian States", grepl("NA", location) ~ "Non Identified", !grepl("Amazonian States|Non Amazonian States|Non Identified", location) ~ "International"))
stage_rate <- data.frame(Stage = names(summary(as.factor(Amazon_lc$stage))),
                         Amazon_speeches = paste0(round(as.numeric(summary(as.factor(Amazon_lc$stage))/861)*100, 2)," %"),
                         All_speeches = paste0(round(as.numeric(summary(as.factor(stages$stage))/6088)*100, 2)," %"))
stage_rate
```

@@

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()
tamazon
```

@@

landscape(f_speaker, margin=NULL)
```

@@

The tabel 2 (above) and table 3(below) portray the result of a structural topic model. We utilize LDA (Latent Dirichlet allocation), which is a model for estimating the mixture of words associated with a topic (cluster) and determining the probabilities of topics in each mention, we start to see some interesting patterns (for more information on LDA and R approaches [here](https://www.tidytextmining.com/index.html)). We set the number of clusters to be found (topics) at 10, randomly. We, then, select the top ten words coded for each topic.

@@

Once the clusters are extracted, we can indicate, probabilistically, how presidents' discourse fall within a topic, that's what the second table does for the 2 most salient topics per speakier.We see that, for example, Lula's discourses about the Amazon often is clustered with themes of development and of deforestation. At the same time, Bolsonaro's discourses about the Amazon often is clustered with themes of integration, energy, and the military forces. Sarney engaged with themes of preservation, while Cardoso engages with themes related to energy and integration. Others as Rousseff and Temer appear to engage with themes of geography, politics, energy; while Collor an Itamar appear to engage with themes of security, development, and integration.

@@

```{r narratives, echo=FALSE, warning=FALSE, message=FALSE, echo=FALSE}

@@

>>>>

---
title: "Amazonian Narratives in Presidential Discourse (1985 to 2020)"
author:
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
date: "11/11/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

@@

Preserving the Amazon has rarely been the top priority of Brazilian federal governments. Though deforestation rates began to steadily increase in 2012 (see Carvalho et al. 2019), after a decade of decline, they spiked in 2019 after Bolsonaro took power (INPE 2021). This is no surprise as the current federal government is weakening the foundations of the Brazilian environmental governance system through budgetary cuts, presidential appointments, and dissociating government from civil society (see Ferrante and Fearnside 2019). These factors contributed to record-high fires in the Amazon and unprecedented monthly rates of deforestation during the dry seasons of 2019, 2020, and 2021 (INPE 2021). Although policy discontinuities between Bolsonaro's Amazon agenda and previous governments have been documented by academic scholarship and specialized media (see Abessa, Famá, and Buruaem 2019; Spring 2021), we lack empirical accounts of how Amazonian discourses changed. Pokorny et. al (2021, p. 2208), for example, argue that Bolsonaro’s Amazonian discourses shifted drastically in relation to earlier governments, giving rise to a “populist neoliberal” phase of discourse and policy developments in the region. However, the authors rely on a few disconnect quotes from Bolsonaro about the Amazon to exemplify phases defined by policy discontinuities. *How, then, have Amazonian presidential discourses changed in time?*

@@

*Amazonian narratives constitute discourses that connect the region to specific imaginaries of the future and past*. We argue that they touch on shared meanings about the Amazon that are available to the speaker as part of the larger social history of the region and country. To identify a few possible narratives, we utilize environmental historiography about the Amazon. We build Amazonian narratives as ideal types, that is, unified analytical constructs portraying a ‘pure’ version of a phenomenon (Weber 1949). In that sense, they rarely exist in its purest form. It serves the purpose of allocating empirical observations within a range of possibilities. The ideal types also help us deal with statements that perform different narratives. We broadly understand political discourses as performances seeking to appeal to diverse audiences across multiple political stages to project a definition of the situation (Goffman 1956, p. 3). As political stages, audiences, and time change, so do narratives. As with broader political discourses, narratives can appear together, not necessarily in consistent ways in time or across stages. As we show in section 3, our methodological strategy utilizing allows us to allocate statements within one or more ideal types, providing a more sophisticated portrayal of political discourse.

@@

  The developmentalist narrative advances the view that the Brazilian Amazon needs to be developed and modernized. This includes expanding the agricultural frontier through incentives, creating infrastructure as roads and dams, and fostering industries through tax-free zones. It has its origins in the policies of the 1930s that attempted to modernize the recently consolidated region (Hecht and Cockburn, 1990), but were extended throughout the 20th century. It also captures arguments of the 20th century that defended the integration of indigenous populations into the nation via assimilation and inclusion in economic life. This narrative employs imaginaries of underdevelopment and economic needs in the Amazon while, at the same time, focus on the unexplored economic potential of the region and forest.
\newpage
*Conservationism*

@@

# 4 Exploratory analysis: Amazonian narratives along time

@@

# Remove stop words and punctuation
Amazon_speeches$text <- tm::removePunctuation(Amazon_speeches$text)
Amazon_speeches$text <- tm::removeWords(Amazon_speeches$text,
                                        c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre", "roraima",
                                          "manaus", "rondônia", "ministros", "luiz", "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "rio grande do sul", "rio grande do norte", "sao paulo", "então", "pouco", "pode", "sul",
                                          "norte", "mato grosso","outro", "onde", "discurso", "parte", "anos", "cada", "todo", 
                                          "porto velho", "visite site", "site secretaria", "minas gerais", "lula silva", 
                                          "inácio lula", "amapa", "bahia", "é", "lá", "et", "in", "httpwww", "vez", "vezes", "entao", "coisa", "ainda", "sobre", "apenas", "coisa",
                                          "jose", "tudo", "toda", "mil", "sempre", "assim", "estados", "forma", "pais", "programa", "primeiro"))

@@

```{r location, echo=FALSE, warning=FALSE, message=FALSE}
# Location was added with poldis::extraxt_location(), still needs to be improved...
# To see the variable summary, please run: `summary(as.factor(Amazon_speeches$location))`
# Let's code the Amazonian states, the other Brazilian states, and speeches delivered outside Brazil.
Amazon_lc <- mutate(Amazon_speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian_States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non_Amazonian_States", grepl("NA", location) ~ "Non_Identified", !grepl("Amazonian_States|Non_Amazonian_States|Non_Identified", location) ~ "International"))
# Just a comparison with the proportion of speeches per stage on the broader dataset
stages <- mutate(BR_Presidential_Speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non Amazonian States", grepl("NA", location) ~ "Non Identified", !grepl("Amazonian States|Non Amazonian States|Non Identified", location) ~ "International"))
stage_rate <- data.frame(Stage = names(summary(as.factor(Amazon_lc$stage))),
                         Amazon_speeches = paste0(round(as.numeric(summary(as.factor(Amazon_lc$stage))/861)*100, 2)," %"),
                         All_speeches = paste0(round(as.numeric(summary(as.factor(stages$stage))/6088)*100, 2)," %"))
knitr::kable(stage_rate, caption= "Speeches by stage")
```

@@

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()

@@

landscape(f_speaker, margin=NULL)

@@

  The table 2 (above) and table 4(below) portray the result of a structural topic model. We utilize LDA (Latent Dirichlet allocation), which is a model for estimating the mixture of words associated with a topic (cluster) and determining the probabilities of topics in each mention, we start to see some interesting patterns (for more information on LDA and R approaches [here](https://www.tidytextmining.com/index.html)). We set the number of clusters to be found (topics) at 10, randomly. We, then, select the top ten words coded for each topic.

@@

  Once the clusters are extracted, we can indicate, probabilistically, how presidents' discourse fall within a topic, that's what the second table does for the 2 most salient topics per speakier.We see that, for example, Lula's discourses about the Amazon often is clustered with themes of development and of deforestation. At the same time, Bolsonaro's discourses about the Amazon often is clustered with themes of integration, energy, and the military forces. Sarney engaged with themes of preservation, while Cardoso engages with themes related to energy and integration. Others as Rousseff and Temer appear to engage with themes of geography, politics, energy; while Collor an Itamar appear to engage with themes of security, development, and integration.
\newpage

@@

```{r narratives I, echo=FALSE, warning=FALSE, message=FALSE, echo=FALSE}

@@

```{r narratives II, echo=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
